% Context: Why now

\noindent\textbf{Context:} 
The majority ($\sim85$\%) of the matter in the universe is composed of dark matter, a mysterious particle that does not interact via the electromagnetic force yet does interact with all other matter via the gravitational force.
Many direct detection experiments have been devoted to finding interactions of dark matter with baryonic matter via the weak force.
To date only tentative and highly controversial evidence for such interactions has been found.
While such direct detection experiments have ruled out the possibility that dark matter interacts with baryonic matter via a strong scale force, it is still possible that dark matter interacts with itself via a strong scale force and has a self-scattering cross-section of $\sim0.5$\,cm$^2$\,g$^{-1}$.
In fact such a strong scale scattering force could resolve several outstanding astronomical mysteries: a discrepancy between the cuspy density profiles seen in $\Lambda$CDM simulations and the cored density profiles observed in low surface brightness galaxies, dwarf spheroidal galaxies, and galaxy clusters, as well as the discrepancy between the significant number of massive Milky Way dwarf spheroidal halos predicted by $\Lambda$CDM and the dearth of observed Milky Way dwarf spheroidal halos.
 

% Need: What you want is not what you have

\noindent\textbf{Need:} 
While such observations are in conflict with $\Lambda$CDM and suggest that dark matter may self-scatter, each suffers from a \emph{baryonic degeneracy}, where the observations might be explained by various baryonic processes (e.g., AGN or supernove feedback, XXX etc.) rather than self-interacting dark matter (SIDM).
In fact the important scales of these observations often coincide with baryonic scales (e.g., the core size in clusters is approximately the size of the brightest cluster galaxy).
What is needed is a probe of SIDM where the expected effect cannot be replicated by the same processes responsible for the baryonic degeneracy in the aforementioned probes.
Merging galaxy clusters are such a probe.
During the merging process the effectively collisionless galaxies ($\sim$2\% of the cluster mass) become dissociated from the collisional intracluster gas ($\sim$15\% of the cluster mass).
The gas strongly interacts during the merger and becomes pancaked at the point of collision.
If dark matter lags behind the effectively collisionless galaxies then this is clear evidence that dark matter self-interacts.
The expected galaxy-dark matter offset is typically $>$25\,kpc (for cross-sections that would explain the other aforementioned mysteries), this is considerably larger than the scales of that are plagued by the baryonic degeneracies.

%Task: What I did to address the need
\noindent\textbf{Task:} 
To test whether dark matter self-interacts we have carried out a comprehensive survey of the dissociative merging galaxy cluster DLSCL J0916.2+2951 (also known as the Musket Ball Cluster).
This survey includes photometric and spectroscopic observations to quantify the position and velocity of the cluster galaxies,
weak gravitational lensing observations to map and weigh the mass (i.e. dark matter) of the cluster,
Sunyaev-Zel'dovich effect and X-ray observations to map and quantify the intracluster gas,
and finally radio observations to search for associated radio relics that could help constrain the properties of the merger.
Using this information in conjunction with a Monte Carlo analysis model I quantify the dynamic properties of the merger, necessary to properly interpret constraints on the SIDM cross-section.
I compare the locations of the galaxies, dark matter and gas to constrain the SIDM cross-section.
This dissertation presents this work.

% Findings: What I found doing the task
\noindent\textbf{Findings:} 
We find that the Musket Ball is a modest merger with total mass of  $4.8^{+3.2}_{-1.5}\times 10^{14}$M$_\sun$.
However, my dynamic analysis shows that the Musket Ball is being observed $1.1^{+1.3}_{-0.4}$\,Gyr after first pass through and is much further progressed in its merger process than previously identified dissociative mergers ($3.4^{+3.8}_{-1.4}$ times further progressed that the Bullet Cluster). 
By observing that the dark matter is significantly offset from the gas we are able to place an upper limit on the dark matter cross-section of $\sigma_{\rm SIDM}m^{-1}_{\rm DM} <$8\,cm$^2$\,g$^{-1}$.
We find an offset of XXX between the galaxies and dark matter in the southern subcluster.


% Conclusion: What these findings mean to me
\noindent\textbf{Conclusion:}
This offset suggests that dark matter self-interacts with XXX\% confidence, and appears to be consistent with existing SIDM constraints.
While this offset is significant, it is not significant enough to claim that dark matter self-interacts.

% Perspectives: What we should do next
\noindent\textbf{Perspectives:}
The galaxy-dark matter offset measurement is dominated by random errors in each cluster.
Thus measuring this offset in other dissociative mergers holds the promise of reducing our uncertainty and enabling us to: 1) state confidently that dark matter self-interacts via a new dark sector force, or 2) constrain the dark matter cross-section to such a degree that SIDM cannot explain the aforementioned mysteries.
To this end we have established the merging cluster collaboration to observe and simulate an ensemble of dissociative merging cluster.
We are currently in the process of analyzing six dissociative mergers with existing data, and carrying out multi-wavelength observations of a new sample of 15 radio relic identified dissociative mergers.